---
title: "Comparing accounts of formant normalization against US English listeners’ vowel perception"
shorttitle: "Comparing normalization against perception"

author:
  - name: Anna Persson
    affiliation: Swedish Language and Multilingualism, Stockholm University
    # etal: Persson # First author's last name.
    email: anna.persson\@su.se # Indicates corresponding Author
    institution: Department of Swedish Language and Multilingualism
    street: Stockholm University
    city: Stockholm
    zip: SE-106 91
    country: Sweden
  - name: Santiago Barreda
    affiliation: Linguistics, University of California, Davis
  - name: T. Florian Jaeger
    affiliation: Brain and Cognitive Sciences, Data Science, University of Rochester

# The following prints the current date (optional, set to false to remove).
date: true
classoption:
  - preprint # Set to "reprint" or "preprint" 
# The reprint style typesets the manuscript with two columns and 10pt font size.  
#  - NumberedRefs # Uncomment to use numbered references rather than the default author-year style.
#  - trackchanges # Uncomment to track changes in the manuscript. (Use with preprint option).
#  - 12pt # Uncomment this to check if you exceeded the 12 page limit. (Use with reprint option).
#  - TurnOnLineNumbers # Uncomment to add line numbers in reprint. (Use with reprint.)

preprint_notice: "Persson, Barreda, & Jaeger, JASA"

abstract:
# Put your abstract here. Abstracts are limited to 200 words for
# regular articles and 100 words for Letters to the Editor. Please no
# personal pronouns, also please do not use the words ``new'' and/or
# ``novel'' in the abstract. An article usually includes an abstract, a
# concise summary of the work covered at length in the main body of the
# article.
  Human speech perception tends to achieve robust speech recognition, despite substantial cross-talker variability. Believed to be critical to this ability are auditory normalization mechanisms whereby listeners adapt to individual differences in vocal tract physiology in vowel perception. This study asks what types of computations are involved in such normalization. Two 8-way alternative forced-choice experiments assessed L1 listeners' categorizations across the entire US English vowel space---both for unaltered and for synthesized stimuli. Listeners' responses in these experiments were compared against the predictions of twenty influential normalization accounts that differ starkly in the inference and memory capacities they imply for speech perception. Listeners' responses were best explained by *extrinsic* normalization accounts, suggesting that listeners learn and store distributional properties of talkers' speech. Of the extrinsic accounts, it was the *computationally least complex* variants that best fit listeners' responses, using a single parameter. These findings have consequences for any research that aims to investigate the perceptual, social, and linguistic information of vowel productions. This includes research in phonetics and phonology, sociolinguistics, and language acquisition. In these fields, it remains common to employ normalization accounts that the present study confirms to be suboptimal models of human perception (e.g., Lobanov normalization). 

keywords          : "speech perception; normalization; vowels; formants; ideal observers"
bibliography      : ["latex-stuff/library.bib", "latex-stuff/r-references.bib"]

output:
  bookdown::pdf_book:
    base_format: rticles::jasa_article
    latex_engine: xelatex
    includes:
      in_header: latex-stuff/header.tex
    keep_tex: yes
# The Author-Year style is the default. If you want to use the numeric style,
# use "jasanum2.bst" instead.
biblio-style: jasaauthyear2.bst
---

```{r source-preamble, include=FALSE}
source("preamble.R")
```

```{r global-options, include=FALSE}
INCLUDE_TODO = F        # switch on/off to get to do list.
INCLUDE_SI = T          # switch on/off to get SI.
RESET_MODELS = F        # switch on/off whether model optimization etc. is re-run

# Use fig.process to rename figure files generated by R so that they comply with
# the journal naming policies. Remember to name plot chunks with the format
# "Figure1", "Figure2", etc... You will find the figure files in
# `[...]_files/figure-latex/`.

# The only figure formats allowed for submission are the following: `.pdf`, `.ps`, `.eps`, or `.jpg`.
# Figure files must be named in this fashion:
# `Figure\#.xxx`, where `\#` is the figure number and `xxx` is the file format
# (`Figure1.eps`, `Figure2.jpg`, `Figure3a.ps`, `Figure3b.ps`, etc).

knitr::opts_chunk$set(
  out.width="\\reprintcolumnwidth",
  fig.process = function(x) {
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

# The JASA LaTeX class depends on the LaTeX package "revtex4-2". If you are
# using a full LaTeX distro, you should make sure "revtex4-2" is installed, and
# install it if not. 
```

\setcounter{secnumdepth}{5}
\setcounter{page}{1}
\setcounter{section}{0}

```{r sec-intro, child="section-introduction.Rmd", eval=TRUE}
```

```{r sec-experiments, child="section-experiments.Rmd", eval=TRUE}
```

```{r sec-study1, child="section-computational-comparison.Rmd", eval=TRUE}
```

```{r sec-study2, child="section-general-discussion.Rmd", eval=TRUE}
```

\begin{acknowledgments}
Earlier versions of this work were presented at 2023 ASA meeting, ExLing 2022, at the Department of Computational Linguistics at the University of Zürich and at the Department of Swedish language and multilingualism at Stockholm University. We are grateful to OMITTED FOR REVIEW.
\end{acknowledgments}
<!-- Maryann Tan, Chigusa Kurumada, and Xin Xie for feedback on this work. We thank Leslie Li and Xin Xie for sharing their database of L1-US English *hVd* productions, and the JASA copy editing staff for help with the Latex formatting. This work was partially funded by grants to AP from Kungliga Vetenskapsakademien, Kungliga Vitterhetsakademien, and the Department of Swedish Language and Multilingualism at Stockholm University, as well as grants to TFJ by the Helge Ax:son Johnson foundation, the Stockholm University Board of Human Science (Funding for Strategic Investments), and the Stockholm University Faculty of Humanities’ Research School (Kvalitetssäkrande medel grant). -->


# Author Contributions {-}
AP designed the experiments and collected the data, with input from TFJ. TFJ programmed the experiments with input from AP. AP analyzed the experiments, with input from TFJ. AP and TFJ wrote the code to implement and fit the normalization models, with input from SB. AP developed the visualizations within input from SB and TFJ. AP wrote the first draft of the manuscript with edits by SB and TFJ.


# Author Declarations {-}
## Conflict of Interest {-}
The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

## Ethics approval {-}
This study was reviewed and approved Research Subjects Review Board (RSRB) of the University of Rochester (STUDY00000417) under the OHSP and UR policies, and in accordance with Federal regulation 45 CFR 46 under the university’s Federal-wide Assurance (FWA00009386).

```{r, child="SI.Rmd", eval= if (INCLUDE_SI) TRUE else FALSE}
```

\newpage{}

# References {#sec:references}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>

\nocite{barreda-nearey2018}
\nocite{goldinger1996}
\nocite{hay2017}
\nocite{hay2019}
\nocite{johnson1999}
\nocite{kleinschmidt-jaeger2016}
\nocite{mcgowan2015}
\nocite{shannon1948}
\nocite{sumner2011}
\nocite{walker-hay2011}


\endgroup
