---
title: "Comparing accounts of formant normalization against US English listeners’ vowel perception"
shorttitle: "Comparing normalization against perception"

author:
  - name: Anna Persson
    affiliation: Swedish Language and Multilingualism, Stockholm University, Stockholm, SE-106 91, Sweden
    # etal: Persson # First author's last name.
    email: anna.persson\@su.se # Indicates corresponding Author
    institution: Department of Swedish Language and Multilingualism
    street: Stockholm University
    city: Stockholm
    zip: SE-106 91
    country: Sweden
  - name: Santiago Barreda
    affiliation: Linguistics, University of California, Davis
  - name: T. Florian Jaeger
    affiliation: Brain and Cognitive Sciences, Data Science, University of Rochester

# The following prints the current date (optional, set to false to remove).
date: true
classoption:
  - preprint # Set to "reprint" or "preprint"
# The reprint style typesets the manuscript with two columns and 10pt font size.  
#  - NumberedRefs # Uncomment to use numbered references rather than the default author-year style.
#  - trackchanges # Uncomment to track changes in the manuscript. (Use with preprint option).
#  - 12pt # Uncomment this to check if you exceeded the 12 page limit. (Use with reprint option).
#  - TurnOnLineNumbers # Uncomment to add line numbers in reprint. (Use with reprint.)

preprint_notice: "Persson, Barreda, & Jaeger, JASA"

abstract:
# Put your abstract here. Abstracts are limited to 200 words for
# regular articles and 100 words for Letters to the Editor. Please no
# personal pronouns, also please do not use the words ``new'' and/or
# ``novel'' in the abstract. An article usually includes an abstract, a
# concise summary of the work covered at length in the main body of the
# article.
  Human speech recognition tends to be robust, despite substantial cross-talker variability. Believed to be critical to this ability are auditory normalization mechanisms whereby listeners adapt to individual differences in vocal tract physiology. This study investigates the computations involved in such normalization. Two 8-way alternative forced-choice experiments assessed L1 listeners' categorizations across the entire US English vowel space---both for unaltered and for synthesized stimuli. Listeners' responses in these experiments were compared against the predictions of twenty influential normalization accounts that differ starkly in the inference and memory capacities they imply for speech perception. This includes variants of *estimation-free* transformations into psycho-acoustic spaces, *intrinsic* normalizations relative to concurrent acoustic properties, and *extrinsic* normalizations relative to talker-specific statistics. Listeners' responses were best explained by extrinsic normalization, suggesting that listeners learn and store distributional properties of talkers' speech. Specifically, *computationally simple* (single-parameter) extrinsic normalization best fit listeners' responses. This simple extrinsic normalization also clearly outperformed Lobanov normalization---a computationally more complex account that remains popular in research on phonetics and phonology, sociolinguistics, typology, and language acquisition.
  
keywords          : "speech perception; normalization; vowels; formants; ideal observers"
bibliography      : ["latex-stuff/library.bib", "latex-stuff/r-references.bib"]

output:
  bookdown::pdf_book:
    base_format: rticles::jasa_article
    latex_engine: pdflatex
    includes:
      in_header: latex-stuff/header.tex
    keep_tex: yes
# The Author-Year style is the default. If you want to use the numeric style,
# use "jasanum2.bst" instead.
biblio-style: jasaauthyear2.bst
---

```{r source-preamble, include=FALSE}
source("preamble.R")
```

```{r global-options, include=FALSE}
# % In order to compile the document without the SI, while still having all 
# % cross-reference to the SI, we're using the xr package:
# % 1) Knit the document with INCLUDE_SI = T.
# % 2) Rename the .aux file to "manuscript-with-SI.aux"
# % 3) Reknit the document with INCLUDE_SI = F
# % 
# % For this to work:
# % a) the xr latex package needs to be loaded as part of the preamble of the 
# %    document you're knitting: \usepackage{xr}
# % b) "manuscript-with-SI" needs to be part of the preamble of the document
# %    you're knitting: \externaldocument{manuscript-with-SI}
# % c) options(tinytex.clean = FALSE) needs to be specified in the R code, so 
# %    that the aux and bibliography files are kept after knitting.
INCLUDE_SI = F          # switch on/off to get SI.
INCLUDE_TODO = F        # switch on/off to get to do list.
RESET_MODELS = F        # switch on/off whether model optimization etc. is re-run
```

\setcounter{secnumdepth}{5}
\setcounter{page}{1}
\setcounter{section}{0}

```{r sec-intro, child="section-introduction.Rmd", eval=TRUE}
```

```{r sec-experiments, child="section-experiments.Rmd", eval=TRUE}
```

```{r sec-study1, child="section-computational-comparison.Rmd", eval=TRUE}
```

```{r sec-study2, child="section-general-discussion.Rmd", eval=TRUE}
```

# Supplementary Material {-}
See supplementary material at [https://osf.io/zemwn/](https://osf.io/zemwn/) for more details on participant and experiment data, on the vowel database used, and on the computational models presented in the paper. The supplementary material also contains additional auxiliary analyses, including models trained on different subsets of the data, and on additional cues besides F1-F2.

\begin{acknowledgments}
Earlier versions of this work were presented at 2023 ASA meeting, ExLing 2022, at the Department of Computational Linguistics at the University of Zürich and at the Department of Swedish language and multilingualism at Stockholm University. We are grateful to Maryann Tan, Chigusa Kurumada, and Xin Xie for feedback on this work. We thank Travis Wade for clarifications on the synthesis procedure used in his study. We thank Leslie Li and Xin Xie for sharing their database of L1-US English *hVd* productions, and the JASA copy editing staff for help with the Latex formatting. This work was partially funded by grants to AP from Kungliga Vetenskapsakademien, Kungliga Vitterhetsakademien, and the Department of Swedish Language and Multilingualism at Stockholm University, as well as grants to TFJ by the Helge Ax:son Johnson foundation, the Stockholm University Board of Human Science (Funding for Strategic Investments), and the Stockholm University Faculty of Humanities’ Research School (Kvalitetssäkrande medel grant).
\end{acknowledgments}


# Author Contributions {-}
AP designed the experiments and collected the data, with input from TFJ. TFJ programmed the experiments with input from AP. AP analyzed the experiments, with input from TFJ. AP and TFJ wrote the code to implement and fit the normalization models, with input from SB. AP developed the visualizations within input from SB and TFJ. AP wrote the first draft of the manuscript with edits by SB and TFJ.


# Author Declarations {-}
## Conflict of Interest {-}
The authors have no conflicts to disclose.

## Ethics approval {-}
This study was reviewed and approved Research Subjects Review Board (RSRB) of the University of Rochester (STUDY00000417) under the OHSP and UR policies, and in accordance with Federal regulation 45 CFR 46 under the university’s Federal-wide Assurance (FWA00009386).

```{r, child="SI.Rmd", eval= if (INCLUDE_SI) TRUE else FALSE}
```

\newpage{}

# References {#sec:references}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>

\nocite{balzano1982}
\nocite{barreda-nearey2018}
\nocite{goldinger1996}
\nocite{hay2017}
\nocite{hay2019}
\nocite{johnson1999}
\nocite{kleinschmidt-jaeger2016}
\nocite{mcgowan2015}
\nocite{shankweiler1978}
\nocite{shannon1948}
\nocite{strange-jenkins2012}
\nocite{sumner2011}
\nocite{walker-hay2011}
\nocite{zwicker1957}
\nocite{zwicker-terhardt1980}

\endgroup
